---
title: "Assignment 2 Appendix"
output: pdf_document
---

```{r}
library(survival)
library(tidyverse)
knitr::opts_chunk$set(fig.width=3, fig.height=3) 
dat = read.csv('C:/Users/krss9/Desktop/FS21/Stats 504/Assignment2/diabeticVision.csv')

#trt = 0 means that eye did not get a treatment. We will change those laser types to none.
#for(i in 1:nrow(dat))
#if(dat$trt[i] == 1){
#  dat$laser[i] ='none'
#}
#dat %>% filter(dat$laser == 'none') %>% head()
```
```{r}
dat['trt2'] = 0
dat[dat$trt == 0, 'trt2'] = 'control'
dat[dat$trt == 1, 'trt2'] = 'treated'
head(dat)
```

```{r}
sd(dat$futime)
```
```{r}
sum(dat$laser == 'xenon')
sum(dat$laser == 'argon')
sum(dat$eye == 'left')
sum(dat$eye == 'right')
sum(dat$status == '0')
sum(dat$status == '1')
sum(dat$type == 'adult')
sum(dat$type == 'juvenile')
sum(dat$trt == '0')
```

#Goals
### 1. Determine the efficacy of treatment type on visual acuity 
### 2. Quantify the improvement between eyes by treatment type


### 3. Understand the potential impact that age at diagnosis have on visual acuity
### 4. understand the potential impact that clinical risk of diabetic retinopathy have on visual acuity

### 0. Overall survival
```{r}
survobj <- with(dat, Surv(futime, status))
fit0 <- survfit(survobj~1, data=dat)
```

```{r}
plot(fit0, xlab="Time at risk of loss of vision in Months",
ylab="% not lost vision", yscale=100,
main ="Survival Distribution (Overall)")
```

### 1. Survival distribution by laser
```{r}
#dat['inter'] = dat['laser'] * dat['trt']
dat$laser_trt = paste(as.factor(dat$trt2),as.factor(dat$laser) )
fitr <- survfit(survobj~ laser_trt, data=dat)
plot(fitr, xlab="Time at risk of loss of vision in Months",
ylab="% not lost vision", yscale=100,
main="Survival Distribution by laser",
col = c('red', 'blue','green','black'))
legend('bottomleft', legend=levels(as.factor(dat$laser_trt)), col = c('red', 'blue','green', 'black'), lty=1)
```
### 2. Quantatitive analysis of the survival distribution by laser
```{r}
survdiff(survobj~laser, data=dat)
```


### 3. Survival Distrubution by age - graph

```{r}
summary(dat$age)
```


```{r}
dat$agedist <- cut(dat$age, breaks=c(0,18,30,100))
fitc <- survfit(survobj~agedist, data=dat)
plot(fitc,xlab="Time at risk of loss of vision in Months",
ylab="% not lost vision", yscale=100,
main="Survival Distribution by age",
col = c('orange', 'yellow', 'green'))
legend('bottomleft', legend=levels(as.factor(dat$agedist)), col = c('orange', 'yellow', 'green'), lty=1)
```

### Survival distribution by clinical risk - graph

```{r}
summary(dat$risk)
```

```{r}
dat$riskdist <- cut(dat$risk, breaks=c(5,8,10,12))
fitr <- survfit(survobj~riskdist, data=dat)
plot(fitr,xlab="Time at risk of loss of vision in Months",
ylab="% not lost vision", yscale=100,
main="Survival Distribution by clinical risk of diabetic retinopathy",
col = c('yellow', 'green', 'purple'))
legend('bottomleft', legend=levels(as.factor(dat$riskdist)), col = c('yellow', 'green', 'purple'), lty=1)
```

### More Quantatitive analysis

```{r}
model1 = coxph(survobj~laser * trt + age + risk , data=dat)
summary(model1)
```
```{r}
head(dat)
```
```{r}
#dat$laser = relevel(as.factor(dat$laser), ref='argon')
#coxph(survobj~laser, data=dat)
```



```{r}
#coxph(survobj~agedist, data=dat)
```
```{r}
#coxph(survobj~riskdist, data=dat)
```

### Check the interactino between variables

```{r}
#test.dat1 <- cox.zph(coxph(survobj~laser*age*risk, data=dat))
#test.dat1
```
```{r}
#test.dat2 <- cox.zph(coxph(survobj~laser*risk, data=dat))
#test.dat2
```
