---
title: "Stats_551_final_project"
author: "Anil, Heather, Youngwoo"
date: "4/19/2022"
output:
  rmarkdown::html_document:
    toc: TRUE
    toc_float:
      collapsed: True
      smooth_scroll: True
    theme: flatly
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
responses <- read_csv("https://raw.githubusercontent.com/OxCGRT/covid-policy-tracker/master/data/OxCGRT_latest.csv")
library(sigmoid)
library(clipr)
library(dplyr)
period <- 1
```

# Introduction

Covid-19, also named as 2019-nCoV, 2019 novel coronavirus by the World Health Organization, is in charge of the current outbreak of pneumonia which began in December 2019 in China.  . Compared with the notorious outbreaks of Ebola or SARS, COVID-19 showed relatively low mortality rate but faster spread rate via mildly symptomatic cases. In response to the spread of the corona, the public health and government promoted various policies such as social distancing requirements, masks requirements, stay-at-home orders, or travel restrictions to limit or prevent person-to-person interactions. The objective of our project is to conduct an analysis of the United States, “country 1”, and “country 2” with a “Method”.

# Data

```{r}
source("data/get_case_data.R")
```

```{r}
# data
data = cases
rcountries = c("United States")
data <- data %>%
  filter(location %in% rcountries) %>%
  filter(!is.na(new_cases)) %>%
  group_by(location) %>%
  mutate(recoveries = generate_recovery(new_cases)) %>%
  mutate(currently_infected = cumsum(new_cases) - cumsum(recoveries)) %>%
  select(location, currently_infected, susceptible_estimate, new_cases, recoveries, date) %>%
  rename(I=currently_infected, S=susceptible_estimate, Infection=new_cases, Recovery=recoveries, day=date)  %>%
  ungroup()
```

```{r}
#Infection in United States
ggplot(data, aes(x = day, y = I, color = location)) +
  geom_line()
```


```{r}
head(cases[cases$location=="United States",])
N<-332915074
```

```{r}
rcon1_data = cases_orig %>% filter(location == "United States")
```

```{r}
#Fully vaccination in United States
plot(rcon1_data$people_fully_vaccinated_per_hundred)
```

```{r}
head(cases_orig %>% filter(location == "United States") %>% filter(people_fully_vaccinated_per_hundred > 10))
```

```{r}
# Mask policy in United States
 cases_2 = cases[,c("location", "date", "new_cases", "total_cases","population")]
 responses_2 = responses[,c("CountryName", "Date", "H6_Facial Coverings")]
 colnames(responses_2) = c("location", "date", "mask")
 cases_2$date = 10000 * year(cases_2$date) + 100 * month(cases_2$date) + day(cases_2$date)
 total <- merge(cases_2,responses_2, by =c("location","date"))
 total <- drop_na(total)
 total$date <- as.Date(as.character(total$date), format = "%Y%m%d")
 total = total[!duplicated(total),]
 total_us = total[total$location=="United States",]

 total_us = total_us %>% group_by(location, date, new_cases, total_cases,population) %>% mutate(mask_1 = mean(mask)) %>% select(-mask) %>% unique() %>% ungroup()
```

```{r}

 ggplot(data=total_us)+
   geom_line(aes(x=date, y=mask_1))
```

```{r}
 df_response = data.frame(total_us[1,])

 for(i in 2:nrow(total_us)){
   if(total_us$mask_1[i] != total_us$mask_1[i-1]){
     df_response[nrow(df_response) + 1,]<- total_us[i,]
   }

 }
 df_response
```

Four criteria for slicing the data.

2020-10-19 : mask restriction (2.5->3)

2021-01-12 : fully vaccinated rate reached 50%

2021-05-23 : mask (3->2.5)

2021-11-01 : Omicron

```{r}
# Five data sets based on the criteria
data_1 = data %>% filter(day<"2020-10-19")
data_2 = data %>% filter(day>"2020-10-18") %>% filter(day<"2021-05-23")
data_3 = data %>% filter(day>"2021-05-22") %>% filter(day<"2021-07-13")
data_4 = data %>% filter(day>"2021-07-12") %>% filter(day<"2021-11-01")
data_5 = data %>% filter(day>"2021-10-31") %>% filter(day<"2022-04-10")
```

```{r}
head(data_1)
```

# Statistical Methods

For the Metropolis - Hasting algorithm, we need the following components: - Likelihood function - Prior distribution - Proposal distribution - Proposal sampler

### Likelihood function:

```{r}
#Probability function for the infection and recovery
prob_inf <- function(beta, s, i, x, N){
  rate = period*beta*s*i/N
  return(dpois(x, rate, log = TRUE))
}

prob_rec <- function(gamma, i, x){
  rate <- period*gamma*i
  return(dpois(x, rate, log = TRUE))
}
```

```{r}
#Likelihood function for infection and recovery
likelihood_infection <- function(data, beta){
  ll <- 0
  days <- data$day
  for (d in days){
    drow <- data[data$day == d, ]
    s <- drow$S
    i <- drow$I
    x_inf <- drow$Infection
    ll <- ll + (prob_inf(beta, s, i, x_inf, N))
  }
  return(ll)
}

likelihood_rec <- function(data, gamma){
  ll <- 0
  days <- data$day
  for (d in days){
    drow <- data[data$day == d, ]
    i <- drow$I
    x_rec <- drow$Recovery
    ll <- ll + (prob_rec(gamma, i, x_rec))
  }
  return(ll)
}
```

```{r}
#Applied data_1 into likelihood function
lapply(seq(0.01, 0.5, 0.01), likelihood_infection, data=data_1) %>% unlist() %>% plot(x=seq(0.01, 0.5, 0.01))
```


### Prior distribution

We will assume that the prior distribution for beta is a gamma with a mode on 0.2 (we know that the data has been generated with beta=0.25)

```{r}
#Prior distribution
variance <- 0.7
mod <- 0.2
shape <- 1/(variance-mod)
rate <- (shape-1)/mod

gamma_test <- rgamma(1000, shape, rate)
plot(density(gamma_test))
```

```{r}
# beta from the prior distribution
prior_beta <- function(theta){
  return(dgamma(theta, shape, rate, log = TRUE))
}
```

```{r}
lapply(seq(0.01, 0.5, 0.01), prior_beta) %>% unlist %>% plot(x=seq(0.01, 0.5, 0.01))
```

### Proposal distribution:

```{r}
#Probability for the proposal distribution
prop_pdf <- function(theta, theta_prev){
  return(dlnorm(theta, log(0.2)+1, sigmoid(1/theta_prev), log = T))
}
```

Proposal sampler:

```{r}
#Sampler for the Proposal distribution
prop_samplr <- function(theta_prev){
   return(rlnorm(1, log(0.2)+1, sigmoid(1/theta_prev)))
}
```

We will also write a posterior function to simplify our task:

```{r}
#Beta for the posterior distribution
posterior_beta <- function(data, beta){
    likelihood <- likelihood_infection(data, beta)
    prior <- prior_beta(beta)
    return(likelihood+prior)
}
```

```{r}
lapply(seq(0.01, 0.5, 0.01), posterior_beta, data=data_1) %>% unlist %>% plot(x=seq(0.01, 0.5, 0.01))
```


### Metropolis-Hasting algorithm

```{r}
# Estimate beta_1 from the Metropolis-Hasting algorithm
set.seed(0193)
reps <- 1000
pb = txtProgressBar(min = 1, max = reps, initial = 1)
beta_list_1 <- c(0.2)
prop_1 <- c(0.2)
a_list_1 <- c(0)
for (i in 2:reps){
  theta_star <- prop_samplr(last(beta_list_1))
  a <- posterior_beta(data_1, theta_star) + prop_pdf(last(beta_list_1), theta_star) -
    posterior_beta(data_1, last(beta_list_1)) - prop_pdf(theta_star, last(beta_list_1))
  a_list_1 <- c(a_list_1, a)
  a <- exp(a)
  if (a > 1) {
    beta_list_1 <- c(beta_list_1, theta_star)
  }
  else {
    coin <- rbinom(1,1,a)
    if (coin == 1){
      beta_list_1 <- c(beta_list_1, theta_star)
    }
    else{
      prop_1 <- c(prop_1, theta_star)
    }
  }
  setTxtProgressBar(pb,i)
}
save(beta_list_1, file = "save_sims/beta_list_1.rds")
```



```{r}
beta_list_1
```
```{r}
d <- density(beta_list_1)
plot(d)
```

```{r}
length(beta_list_1)
```

### data_2

We will repeat the steps we did in the data_1.

```{r}
lapply(seq(0.01, 0.5, 0.01), likelihood_infection, data=data_2) %>% unlist() %>% plot(x=seq(0.01, 0.5, 0.01))
```

```{r}
variance <- 0.7
mod <- 0.2
shape <- 1/(variance-mod)
rate <- (shape-1)/mod

gamma_test <- rgamma(1000, shape, rate)
plot(density(gamma_test))
```

```{r}
lapply(seq(0.01, 0.5, 0.01), prior_beta) %>% unlist %>% plot(x=seq(0.01, 0.5, 0.01))
```

### Proposal distribution:

```{r}
lapply(seq(0.01, 0.5, 0.01), posterior_beta, data=data_2) %>% unlist %>% plot(x=seq(0.01, 0.5, 0.01))
```


### Metropolis-Hasting algorithm

```{r}
set.seed(3846)
reps <- 1000
pb = txtProgressBar(min = 1, max = reps, initial = 1)
beta_list_2 <- c(0.2)
prop_2 <- c(0.2)
a_list_2 <- c(0)
for (i in 2:reps){
  theta_star <- prop_samplr(last(beta_list_2))
  a <- posterior_beta(data_2, theta_star) + prop_pdf(last(beta_list_2), theta_star) -
    posterior_beta(data_2, last(beta_list_2)) - prop_pdf(theta_star, last(beta_list_2))
  a_list_2 <- c(a_list_2, a)
  a <- exp(a)
  if (a > 1) {
    beta_list_2 <- c(beta_list_2, theta_star)
  }
  else {
    coin <- rbinom(1,1,a)
    if (coin == 1){
      beta_list_2 <- c(beta_list_2, theta_star)
    }
    else{
      prop_2 <- c(prop_2, theta_star)
    }
  }
  setTxtProgressBar(pb,i)
}
save(beta_list_2, file = "save_sims/beta_list_2.rds")
```

```{r}
beta_list_2
```
```{r}
d <- density(beta_list_2)
plot(d)
```

```{r}
length(beta_list_2)
```


### data_3

We will repeat the steps we did in the data_1.

```{r}
lapply(seq(0.01, 0.5, 0.01), likelihood_infection, data=data_3) %>% unlist() %>% plot(x=seq(0.01, 0.5, 0.01))
```


```{r}
variance <- 0.7
mod <- 0.2
shape <- 1/(variance-mod)
rate <- (shape-1)/mod

gamma_test <- rgamma(1000, shape, rate)
plot(density(gamma_test))
```

```{r}
lapply(seq(0.01, 0.5, 0.01), prior_beta) %>% unlist %>% plot(x=seq(0.01, 0.5, 0.01))
```

### Proposal distribution:

```{r}
lapply(seq(0.01, 0.5, 0.01), posterior_beta, data=data_3) %>% unlist %>% plot(x=seq(0.01, 0.5, 0.01))
```

### Metropolis-Hasting algorithm

```{r}
set.seed(3994)
reps <- 1000
pb = txtProgressBar(min = 1, max = reps, initial = 1)
beta_list_3 <- c(0.2)
prop_3 <- c(0.2)
a_list_3 <- c(0)
for (i in 2:reps){
  theta_star <- prop_samplr(last(beta_list_3))
  a <- posterior_beta(data_3, theta_star) + prop_pdf(last(beta_list_3), theta_star) -
    posterior_beta(data_3, last(beta_list_3)) - prop_pdf(theta_star, last(beta_list_3))
  a_list_3 <- c(a_list_3, a)
  a <- exp(a)
  if (a > 1) {
    beta_list_3 <- c(beta_list_3, theta_star)
  }
  else {
    coin <- rbinom(1,1,a)
    if (coin == 1){
      beta_list_3 <- c(beta_list_3, theta_star)
    }
    else{
      prop_3 <- c(prop_3, theta_star)
    }
  }
  setTxtProgressBar(pb,i)
}
save(beta_list_3, file = "save_sims/beta_list_3.rds")
```

```{r}
beta_list_3
```
```{r}
d <- density(beta_list_3)
plot(d)
```

```{r}
length(beta_list_3)
```


### data_4

We will repeat the steps we did in the data_1.

```{r}
lapply(seq(0.01, 0.5, 0.01), likelihood_infection, data=data_4) %>% unlist() %>% plot(x=seq(0.01, 0.5, 0.01))
```

```{r}
variance <- 0.7
mod <- 0.2
shape <- 1/(variance-mod)
rate <- (shape-1)/mod

gamma_test <- rgamma(1000, shape, rate)
plot(density(gamma_test))
```

```{r}
lapply(seq(0.01, 0.5, 0.01), prior_beta) %>% unlist %>% plot(x=seq(0.01, 0.5, 0.01))
```

### Proposal distribution:

```{r}
lapply(seq(0.01, 0.5, 0.01), posterior_beta, data=data_4) %>% unlist %>% plot(x=seq(0.01, 0.5, 0.01))
```

### Metropolis-Hasting algorithm

```{r}
set.seed(4535)
reps <- 1000
pb = txtProgressBar(min = 1, max = reps, initial = 1)
beta_list_4 <- c(0.2)
prop_4 <- c(0.2)
a_list_4 <- c(0)
for (i in 2:reps){
  theta_star <- prop_samplr(last(beta_list_4))
  a <- posterior_beta(data_4, theta_star) + prop_pdf(last(beta_list_4), theta_star) -
    posterior_beta(data_4, last(beta_list_4)) - prop_pdf(theta_star, last(beta_list_4))
  a_list_4 <- c(a_list_4, a)
  a <- exp(a)
  if (a > 1) {
    beta_list_4 <- c(beta_list_4, theta_star)
  }
  else {
    coin <- rbinom(1,1,a)
    if (coin == 1){
      beta_list_4 <- c(beta_list_4, theta_star)
    }
    else{
      prop_4 <- c(prop_4, theta_star)
    }
  }
  setTxtProgressBar(pb,i)
}
save(beta_list_4, file = "save_sims/beta_list_4.rds")
```

```{r}
beta_list_4
```
```{r}
d <- density(beta_list_4)
plot(d)
```

```{r}
length(beta_list_4)
```


### data_5

We will repeat the steps we did in the data_1.

```{r}
lapply(seq(0.01, 0.5, 0.01), likelihood_infection, data=data_5) %>% unlist() %>% plot(x=seq(0.01, 0.5, 0.01))
```


```{r}
variance <- 0.7
mod <- 0.2
shape <- 1/(variance-mod)
rate <- (shape-1)/mod

gamma_test <- rgamma(1000, shape, rate)
plot(density(gamma_test))
```

```{r}
lapply(seq(0.01, 0.5, 0.01), prior_beta) %>% unlist %>% plot(x=seq(0.01, 0.5, 0.01))
```

### Proposal distribution:

```{r}
lapply(seq(0.01, 0.5, 0.01), posterior_beta, data=data_5) %>% unlist %>% plot(x=seq(0.01, 0.5, 0.01))
```

### Metropolis-Hasting algorithm

```{r}
set.seed(3562)
reps <- 1000
pb = txtProgressBar(min = 1, max = reps, initial = 1)
beta_list_5 <- c(0.2)
prop_5 <- c(0.2)
a_list_5 <- c(0)
for (i in 2:reps){
  theta_star <- prop_samplr(last(beta_list_5))
  a <- posterior_beta(data_5, theta_star) + prop_pdf(last(beta_list_5), theta_star) -
    posterior_beta(data_5, last(beta_list_5)) - prop_pdf(theta_star, last(beta_list_5))
  a_list_5 <- c(a_list_5, a)
  a <- exp(a)
  if (a > 1) {
    beta_list_5 <- c(beta_list_5, theta_star)
  }
  else {
    coin <- rbinom(1,1,a)
    if (coin == 1){
      beta_list_5 <- c(beta_list_5, theta_star)
    }
    else{
      prop_5 <- c(prop_5, theta_star)
    }
  }
  setTxtProgressBar(pb,i)
}
save(beta_list_5, file = "save_sims/beta_list_5.rds")
```

```{r}
beta_list_5
```
```{r}
d <- density(beta_list_5)
plot(d)
```

```{r}
length(beta_list_5)
```

```{r}
#Beta_list for each five data sets.
beta_list_1
beta_list_2
beta_list_3
beta_list_4
beta_list_5
```

# Recovery Part


For the Metropolis - Hasting algorithm, we need the following components: - Likelihood function - Prior distribution - Proposal distribution - Proposal sampler

### Likelihood function:

```{r}
#Remove the data that contains NA value
data = data %>% filter(day<"2022-04-10")

#Probability function for recovery state
prob_rec <- function(gamma, i, x){
  rate <- period*gamma*i
  return(dpois(x, rate, log = TRUE))
}
```

```{r}
#Likelihood function for the recovery state
likelihood_rec <- function(data, gamma){
  ll <- 0
  days <- data$day
  for (d in days){

    drow <- data[data$day == d, ]
    i <- drow$I
    if(i!=0){
    x_rec <- drow$Recovery
    ll <- ll + (prob_rec(gamma, i, x_rec))
    }
  }
  return(ll)
}
```


```{r}
lapply(seq(0.01, 0.4, 0.01), likelihood_rec, data=data) %>% unlist %>% plot(x=seq(0.01, 0.4, 0.01))
```


### Prior distribution

We will assume that the prior distribution for beta is a gamma with a mode on 0.2 (we know that the data has been generated with beta=0.2)

```{r}
variance <- 0.25
mod <- 0.2
shape <- 1/(variance-mod)
rate <- (shape-1)/mod

gamma_test <- rgamma(1000, shape, rate)
plot(density(gamma_test))
```


```{r}
prior_gamma <- function(theta){
  return(dgamma(theta, shape, rate, log = TRUE))
}
```


```{r}
lapply(seq(0.01, 0.4, 0.01), prior_gamma) %>% unlist %>% plot(x=seq(0.01, 0.4, 0.01))
```

### Proposal distribution

```{r}
prop_pdf <- function(theta, theta_prev){
  return(dlnorm(theta, log(0.22)+1, 1+sigmoid(1/theta_prev), log = T))
}
```

### Proposal sampler

```{r}
prop_samplr <- function(theta_prev){
   return(rlnorm(1, log(0.22)+1, 1+sigmoid(1/theta_prev)))
}
```

We will also write a posterior function to simplify our task:

```{r}
posterior_gamma <- function(data, gamma){
    likelihood <- likelihood_rec(data, gamma)
    prior <- prior_gamma(gamma)
    return(likelihood+prior)
}
```

```{r}
lapply(seq(0.01, 0.6, 0.01), posterior_gamma, data=data) %>% unlist %>% plot(x=seq(0.01, 0.6, 0.01))
```


### Metropolis-Hasting algorithm

```{r}
set.seed(8402)
reps <- 10000
pb = txtProgressBar(min = 1, max = reps, initial = 1)
gamma_list <- c(0.2)
prop <- c(0.2)
a_list <- c(0)
for (j in 2:reps){
  theta_star <- prop_samplr(last(gamma_list))
  a <- posterior_gamma(data, theta_star) + prop_pdf(last(gamma_list), theta_star) -
    posterior_gamma(data, last(gamma_list)) - prop_pdf(theta_star, last(gamma_list))
  a_list <- c(a_list, a)
  a <- exp(a)
  if (a > 1) {
    gamma_list <- c(gamma_list, theta_star)
  }
  else {
    coin <- rbinom(1,1,a)
    if (coin == 1){
      gamma_list <- c(gamma_list, theta_star)
    }
    else{
      prop <- c(prop, theta_star)
    }
  }
  setTxtProgressBar(pb,j)
}
save(gamma_list, file = "save_sims/gamma_list.rds")
```

```{r}
gamma_list
```

```{r}
d <- density(gamma_list)
plot(d)
```

```{r}
length(gamma_list)
```

# MCMC diagnostics

```{r}
betas <- list(beta_list_1, beta_list_2, beta_list_3, beta_list_4, beta_list_5)
params <- list(beta_list_1, beta_list_2, beta_list_3, beta_list_4, beta_list_5, gamma_list)
save(params, file = "save_sims/all_params.rds")
lapply(betas, FUN = function(x) mean(x[-1]))

load("./save_sims/all_params.rds")
params <- readRDS("./save_sims/all_params.rds")

```

```{r}

```
