---
title: "MCMC final removal"
author: "Anil, Heather, Youngwoo"
date: "4/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sigmoid)
library(dplyr)
```

```{r}
period <- 1/4
N <- 50000
```

### Load the data

```{r}
data_epidemic_aggregated <- readRDS("./data_epidemic_aggregated.rds")
```

### Sneak a peak into the data

```{r}
head(data_epidemic_aggregated[data_epidemic_aggregated$t>2, ])
```

### How to read the data?

Each row in the data says the following: On a day (day), the number of susceptible and infected people in the population are S and I respectively. On the day the number of infection and recovery events are given by the columns 'Infection' and 'Recovery' respectively.

```{r}
data_epidemic_aggregated
```

____________________________________________________________________________________________________________________________



### Metropolis - Hasting algorithm

For the Metropolis - Hasting algorithm, we need the following components: - Likelihood function - Prior distribution - Proposal distribution - Proposal sampler

Likelihood function:

```{r}
prob_rec <- function(gamma, i, x){
  rate <- period*gamma*i
  return(dpois(x, rate, log = TRUE))
}
```

```{r}
likelihood_rec <- function(data, gamma){
  ll <- 0
  days <- data$day
  for (d in days){
    drow <- data[data$day == d, ]
    i <- drow$I
    x_rec <- drow$Recovery
    ll <- ll + (prob_rec(gamma, i, x_rec))
  }
  return(ll)
}
```

```{r}
d_temp <- data_epidemic_aggregated[data_epidemic_aggregated$day > 90 & data_epidemic_aggregated$day < 110, ]
```

```{r}
d_temp
```

Plotting the likelihood

```{r}
lapply(seq(0.01, 0.4, 0.01), likelihood_rec, data=d_temp) %>% unlist() %>% plot(x=seq(0.01, 0.4, 0.01))
```

Prior distribution: We will assume that the prior distribution for beta is a gamma with a mode on 0.25 (we know that the data has been generated with beta=0.3)

```{r}
variance <- 0.95
mod <- 0.15
shape <- 1/(variance-mod)
rate <- (shape-1)/mod

gamma_test <- rgamma(1000, shape, rate)
plot(density(gamma_test))
```

```{r}
prior_gamma <- function(theta){
  return(dgamma(theta, shape, rate, log = TRUE))
}
```

```{r}
lapply(seq(0.01, 0.9, 0.01), prior_gamma) %>% unlist %>% plot(x=seq(0.01, 0.9, 0.01))
```

Proposal distribution:

```{r}
prop_pdf <- function(theta, theta_prev){
  return(dlnorm(theta, log(0.1)+1, 1+sigmoid(1/theta_prev), log = T))
}
```

Proposal sampler:

```{r}
prop_samplr <- function(theta_prev){
   return(rlnorm(1, log(0.1)+1, 1+sigmoid(1/theta_prev)))
}
```


We will also write a posterior function to simplify our task:

```{r}
posterior_gamma <- function(data, gamma){
    likelihood <- likelihood_rec(data, gamma)
    prior <- prior_gamma(gamma)
    return(likelihood+prior)
}
```

```{r}
lapply(seq(0.01, 0.9, 0.01), posterior_gamma, data=d_temp) %>% unlist %>% plot(x=seq(0.01, 0.9, 0.01))
```

```{r}
data_epidemic_aggregated
```


```{r}
data <- data_epidemic_aggregated[data_epidemic_aggregated$day > 90 & data_epidemic_aggregated$day < 120, ]
```


### Metropolis-Hasting algorithm

```{r}
reps <- 10000
pb = txtProgressBar(min = 1, max = reps, initial = 1)
gamma_list <- c(0.1)
prop <- c(0.1)
a_list <- c(0)
for (i in 2:reps){
  theta_star <- prop_samplr(last(gamma_list))
  a <- posterior_gamma(data, theta_star) + prop_pdf(last(gamma_list), theta_star) - 
    posterior_gamma(data, last(gamma_list)) - prop_pdf(theta_star, last(gamma_list))
  a_list <- c(a_list, a)
  a <- exp(a)
  if (a > 1) {
    gammma_list <- c(gamma_list, theta_star)
  }
  else {
    coin <- rbinom(1,1,a)
    if (coin == 1){
      gamma_list <- c(gamma_list, theta_star)
    }
    else{
      prop <- c(prop, theta_star)
    }
  }
  setTxtProgressBar(pb,i)
}
```

```{r}
gamma_list
```


```{r}
d <- density(gamma_list)
plot(d)
```

```{r}
mean(gamma_list)
```

```{r}
length(gamma_list)/reps
```
